---
title: "Red Wine Analysis"
author: "chayut"
date: "2022-11-01"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
---

# 1. Import Library
```{r}
library(tidyverse)
library(corrplot)
library(DT)
```

# 2. Import Dataset
```{r}
wine <- read.csv("wineQualityReds.csv")
```

# 3. Review Dataset
```{r}
# checking missing values
colSums(is.na(wine))
```

```{r}
head(wine)
```

```{r}
tail(wine)
```

```{r}
names(wine)
```

```{r}
str(wine)
```

```{r}
summary(wine)
```

```{r}
table(wine$quality)
```

# 3. Transforming Quality from an Integer to a Factor
```{r}
wine$quality <- factor(wine$quality, ordered = T)
```

### Creating a new Factored Variable called "Rating"
```{r}
wine$rating <- ifelse(wine$quality < 5, "bad",
                      ifelse(wine$quality < 7, "average", "good"))

# order rating
wine$rating <- ordered(wine$rating,
                       levels = c('bad', 'average', 'good'))
```

```{r}
str(wine)
```

```{r}
table(wine$rating)
```

# 4. Univariate Plots (single variable)
### Plot Quality 
```{r}
ggplot(data = wine, aes(x = quality)) +
  geom_bar(width = 1,
           fill = "#146152",
           color = "white") +
  ggtitle("Quality Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Count")
```

### Plot Rating 
```{r}
ggplot(wine, aes(x = rating)) +
  geom_bar(width = 1,
           fill = "#44803F",
           color = "white") +
  ggtitle("Rating Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Rating") +
  ylab("Count")
```

- Quality is distributed from 3 - 8. 
- Most wine exhibit average quality. 
- Very less percentage of wine is of good quality.


### Plot Fixed Acidity 
```{r}
ggplot(wine, aes(x = fixed.acidity)) +
  geom_bar(fill = "#B4CF66") +
  ggtitle("Fixed Acidity Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Fixed Acidity") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = fixed.acidity)) +
  geom_boxplot(fill = "#B4CF66", color = "red") +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Fixed Acidity") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Fixed Acidity") 
```

```{r}
Fixed_Acidity_1 <- ggplot(wine, aes(x = fixed.acidity)) +
                     geom_bar(fill = "#B4CF66") +
                     ggtitle("Fixed Acidity Distribution") +
                     theme_bw() +
                     theme(plot.title = element_text(hjust = 0.5)) +
                     xlab("Fixed Acidity") +
                     ylab("Count")

Fixed_Acidity_2 <- ggplot(wine, aes(x = 1, y = fixed.acidity)) +
                     geom_boxplot(fill = "#B4CF66", color = "red") +
                     geom_jitter(alpha = 0.3) +
                     coord_flip() +
                     ggtitle("The median of Fixed Acidity") +
                     theme(plot.title = element_text(hjust = 0.5)) +
                     xlab("") +
                     ylab("Fixed Acidity") 

library(gridExtra)

grid.arrange(Fixed_Acidity_1, Fixed_Acidity_2, ncol = 2)
```

- The Fixed acidity value seems to dispaly a normal distribution.
- The distribution of Fixed Acidity is positively skewed. 
- The median is around 8 with high concentration of wines with Fixed Acidity.

### Plot Volatile Acidity
```{r}
ggplot(wine, aes(x = volatile.acidity)) +
  geom_histogram(fill = "#FFEC5C",
                 color = "white") +
  ggtitle("Volatile Acidity Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Volatile Acidity") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = volatile.acidity)) +
  geom_boxplot(fill = "#FFEC5C", color = "red") +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Volatile Acidity") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Volatile Acidity") 
```

- The Volatile acidity value seems to dispaly a bimodal normal distribution with two peaks around 0.4 and 0.6.
- The median is around 0.5.

### Plot Citric Acid 
```{r}
ggplot(wine, aes(x = citric.acid)) +
  geom_histogram(fill = "#024059",
                 color = "white") +
  ggtitle("Citric Acid Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Citric Acid") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = citric.acid)) +
  geom_boxplot(fill = "#024059", color = "red", alpha = 0.5) +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Citric Acid") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Citric Acid") 
```

- The distribution of Citric acid looks strange. 
- Some higher values have no data at all and apart from them, the distribution looks almost rectangular. 
- Maybe there was some error in the data or maybe the data collected was incomplete?
- The median is around 0.25.

### Plot Residual Sugar
```{r}
ggplot(wine, aes(x = residual.sugar)) +
  geom_histogram(binwidth = 1,
                 fill = "#A6874E") +
  scale_x_continuous(breaks = seq(0, 15, by = 1)) +
  ggtitle("Residual Sugar Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Residual Sugar") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = residual.sugar)) +
  geom_boxplot(fill = "#A6874E", color = "red") +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Residual Sugar") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Residual Sugar") 
```

- The histogram is highly left skewed (positively skewed).
- There are very less wine samples with residual sugar more than 8.
- The median is around 2.3.

### Plot Chlorides
```{r}
ggplot(wine, aes(x = chlorides)) +
  geom_histogram(fill = "#30A5BF") +
  ggtitle("Chlorides Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Chlorides") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = chlorides)) +
  geom_boxplot(fill = "#30A5BF", color = "red") +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Chlorides") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Chlorides") 
```

- we see a similar distribution like Residual Sugar.
- The median is around 0.056.

### Plot Free Sulfur Dioxide
```{r}
ggplot(wine, aes(x = free.sulfur.dioxide)) + 
  geom_histogram(binwidth = 1,
                 fill = "#731702",
                 color = "white") +
  scale_x_continuous(breaks = seq(0, 80, by = 10)) +
  ggtitle("Free Sulfur Dioxide Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Free Sulfur Dioxide") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = free.sulfur.dioxide)) +
  geom_boxplot(fill = "#731702", color = "red", alpha = 0.5) +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Free Sulfur Dioxide") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Free Sulfur Dioxide") 
```

- Most of the free sulfur dioxide seems to be between the 1-40.
- there is a high peak at 7.

### Plot Total Sulfur Dioxide
```{r}
ggplot(wine, aes(x = total.sulfur.dioxide)) + 
  geom_histogram(binwidth = 5,
                 fill = "#F27405",
                 color = "white") +
  ggtitle("Free Sulfur Dioxide Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Total Sulfur Dioxide") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = total.sulfur.dioxide)) +
  geom_boxplot(fill = "#F27405", color = "red", alpha = 0.5) +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Total Sulfur Dioxide") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Total Sulfur Dioxide") 
```

- we see a similar distribution like Free Sulfur Dioxide.
- Most of the total sulfur dioxide seems to a have a spread between 0 and 150.

### Plot Density
```{r}
ggplot(wine, aes(x = density)) + 
  geom_histogram(fill = "#486966",
                 color = "white") +
  ggtitle("Density Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Density") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = density)) +
  geom_boxplot(fill = "#486966", color = "red", alpha = 0.5) +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Density") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Density") 
```

- This Variable has almost a perfect Normal Distribution.

### Plot pH
```{r}
ggplot(wine, aes(x = pH)) + 
  geom_histogram(fill = "#BD2A2E",
                 color = "white") +
  ggtitle("Density Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Density") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = pH)) +
  geom_boxplot(fill = "#BD2A2E", color = "red", alpha = 0.5) +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Density") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Density") 
```

- pH has a very Normally distributed shape.

### Plot Sulphates
```{r}
ggplot(wine, aes(x = sulphates)) + 
  geom_histogram(fill = "#0099DD",
                 color = "white") +
  ggtitle("Sulphates Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Sulphates") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = sulphates)) +
  geom_boxplot(fill = "#0099DD", color = "red") +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Sulphates") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Sulphates") 
```

- Sulphates also exhibit a similar long tailed distribution like Chlorides or Free/Total Sulphur Dioxide.
- It has relatively less outliers.

### Plot Alcohol
```{r}
ggplot(wine, aes(x = alcohol)) + 
  geom_histogram(fill = "#2E038C",
                 color = "white") +
  ggtitle("Alcohol Distribution") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Alcohol") +
  ylab("Count")
```

```{r}
ggplot(wine, aes(x = 1, y = alcohol)) +
  geom_boxplot(fill = "#2E038C", color = "red", alpha = 0.5) +
  geom_jitter(alpha = 0.3) +
  coord_flip() +
  ggtitle("The median of Alcohol") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Alcohol") 
```

- Alcohol follows a skewed distribution but the skewness is less than that of Chlorides or Residual Sugars.
- The alcohol content seems to vary from 8 to 14 with most of data around 8-10.

# 5. Summary Analysis of the Univariate Plots (Summary single variable)
### Point of Interest
My main point of interest in this dataset is the ‘quality’. I would like to determine which factors determine the quality of a wine.

### Unique features of the dataset
Citric acid has a unique distribution as compared to the other numeric variables. It almost has a rectangular shape apart from a few outliers. Now if we compare the wine quality distribution, this distribution of Citric Acid is very unexpected and maybe there is even a possibility of incomplete data collection.

### Distribution and Outliers
1. Density and pH seems normally distributed with few outliers.
2. Residual sugar and Chloride seems to have extreme outliers.
3. Fixed and volatile acidity, total and free sulfur dioxides, alcohol and sulphates seem to be long-tailed for the outliers present.
4. Citric acid has large number of zero values. I wonder if this is due to incomplete data entry.

# 6. Bivariate Plots (2 variables)
### Plot Correlation
```{r}
correlation <- cor(wine %>% 
                     select(-X) %>% 
                     select(-rating) %>% 
                     mutate(quality_num = as.numeric(quality)) %>% 
                     select(-quality))

corrplot(corr = correlation,
         number.cex = 0.9, 
         method = "circle", 
         type = "upper",
         tl.cex = 0.8,
         tl.col = "black")

library(psych)

wine_2 <- wine %>% 
            select(-X) %>% 
            select(-rating) %>% 
            mutate(quality_num = as.numeric(quality)) %>% 
            select(-quality)

corPlot(wine_2, cex = 0.7)
```

1. Fixed Acidity
    - It shows positive correlation with citric acid which is true since citric acid is one of the fixed acid. 
    - It also shows positive correlation with density.
    - It also shows significant negative correlation with pH and volatile acidity.

2. Volatile Acidity
    - It is highly negatively correlated with citric acid and quality.

3. Free Sulfur Dioxide 
    - It shows significant positive correlation with total SO2, and very less correlation with sulphates.

4. Density
    - Significant negative correlation is observed with alcohol, acidity (fixed and citric acid) and pH

5. Quality
    - Quality and alcohol is positively correlated along with negative correlation with volatile acidity.

### Table Correlation
```{r}
library(pander)

emphasize.strong.cells(which(abs(correlation) > 0.3 & correlation != 1, arr.ind = TRUE))

pandoc.table(correlation)
```

### Plot Fixed Acidity and Quality
```{r}
wine$quality <- as.factor(wine$quality)
wine$rating <- as.factor(wine$rating)

table(wine$quality)
```

```{r}
ggplot(data = wine, aes(x = quality, y = fixed.acidity)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Fixed Acidity and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Fixed Acidity")
```

- Fixed Acidity has almost no effect on the Quality. 
- The mean and median values of fixed acidity remains almost unchanged with increase in quality.

### Plot Volatile Acidity and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = volatile.acidity)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Volatile Acidity and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Volatile Acidity")
```

- Volatile acid seems to have a negative impact on the quality of the wine. 
- As volatile acid level goes up, the quality of the wine degrades.

### Plot Citric Acid and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = citric.acid)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Citric Acid and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Citric Acid")
```

- Citric acid seems to have a positive correlation with Wine Quality.
- Better wines have higher Citric Acid.

### Plot Residual Sugar and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = residual.sugar)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Residual Sugar and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Residual Sugar")
```

- Residual Sugar almost has no effect on the Quality of the Wine. 
- The mean values for the residual sugar is almost the same for every quality of wine.

### Plot Chlorides and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = chlorides)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Chlorides and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Chlorides")
```

- Even though weakly correlated, from the decrease in median values of the Chlorides with increase in quality, it seems that lower percent of Chloride seems to produce better wines.

### Plot Free Sulfur Dioxide and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = free.sulfur.dioxide)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Free Sulfur Dioxide and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Free Sulfur Dioxide")
```

- We see that too low concentration of Free Sulphur Dioxide produces poor wine and too high concentration results in average wine.

### Plot Total Sulfur Dioxide and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = total.sulfur.dioxide)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Total Sulfur Dioxide and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Total Sulfur Dioxide")
```

- As this is a Subset of Free Sulphur Dioxide, we see a similar pattern here.

### Plot Density and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = density)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Density and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Density")
```

- Better wines seems to have lower densities. But may be it will be wise not to draw any conclusions here. Because there might be a possibility that the low density is due to higher alcohol content which actually is the driving factor for better wines.

### Plot pH and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = pH)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("pH and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("pH")
```

- Better wines seems to have lower densities. But may be it will be wise not to draw any conclusions here. Because there might be a possibility that the low density is due to higher alcohol content which actually is the driving factor for better wines.

### Plot Fixed Acidity in Log Scale and pH
```{r}
ggplot(data = wine, aes(x = fixed.acidity, y = pH)) +
  geom_point(alpha = 0.3) +
  scale_x_log10(breaks = seq(5, 15, 1)) +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Fixed Acidity in Log Scale and pH") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Fixed Acidity in Log Scale") +
  ylab("pH")
```

### Plot Volatile Acidity in Log Scale and pH
```{r}
ggplot(data = wine, aes(x = volatile.acidity, y = pH)) +
  geom_point(alpha = 0.3) +
  scale_x_log10(breaks = seq(0.1, 1, 0.1)) +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Volatile Acidity in Log Scale and pH") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  xlab("Volatile Acidity in Log Scale") +
  ylab("pH") 
```

### Plot Citric Acid in Log Scale and pH
```{r}
ggplot(data = subset(wine, citric.acid > 0), aes(x = citric.acid, y = pH)) +
  geom_point(alpha = 0.3) +
  scale_x_log10() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Citric Acid in Log Scale and pH") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Citric Acid in Log Scale") +
  ylab("pH")
```

- These three plots make us come back to our old question. Recall that we saw for Volatile Acid, pH has a positive correlation. But we know acidity has a negative correlation with pH. So is it possible, that we are seeing a Simpson’s Paradox at play here? Let’s investigate.

### Plot Sulphates and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = sulphates)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Sulphates and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Sulphates")
```

- Even though we see many outliers in the ‘Average’ quality wine, it seems that better wines have a stronger concentration of Sulphates.

### Plot Alcohol and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = alcohol)) +
  geom_boxplot(color = "blue", alpha = 0.6) +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Alcohol")
```

- The correlation is really diffence here. 
- It is pretty evident that better wines have higher Alcohol content in it. 
- But we see a great number of outliers here. 
- So it might be possible that alcohol alone does not contribute to a wine being a good quality one.

### Linear Regression between Alcohol and Quality
```{r}
linear_model <- lm(as.numeric(quality) ~ alcohol, data = wine)

summary(linear_model)

library(broom)
library(DT)

datatable(tidy(linear_model))

datatable(glance(linear_model))
```

$$Quality = -0.12503 + 0.36084 \times Alcohol$$

```{r}
library(Metrics)
# Calculate MSE
mean(linear_model$residuals^2)

mse(as.numeric(wine$quality), predict(linear_model, wine))
```

```{r}
# Calculate RMSE
sqrt(mean(linear_model$residuals^2))

rmse(as.numeric(wine$quality), predict(linear_model, wine))
```

- Based on the value of R squared, we see that Alcohol alone contributes to only about 22.67 % of the Wine quality. 
- So there must be other variables at play here. 

### Plot correlation each variable to the quality of the wine
we will put a correlation test against each variable to the quality of the wine.

```{r}
wine_2 <- wine %>% 
            select(-X) %>% 
            select(-rating) %>% 
            mutate(quality_num = as.numeric(quality)) %>% 
            select(-quality)

corPlot(wine_2, cex = 0.7)
```

```{r}
simple_cor_test <- function(x, y) {
  return(cor.test(x, as.numeric(y))$estimate)
}

correlations_2 <- c(simple_cor_test(wine$fixed.acidity, wine$quality),
                    simple_cor_test(wine$volatile.acidity, wine$quality),
                    simple_cor_test(wine$citric.acid, wine$quality),
                    simple_cor_test(log10(wine$residual.sugar), wine$quality),
                    simple_cor_test(log10(wine$chlorides), wine$quality),
                    simple_cor_test(wine$free.sulfur.dioxide, wine$quality),
                    simple_cor_test(wine$total.sulfur.dioxide, wine$quality),
                    simple_cor_test(wine$density, wine$quality),
                    simple_cor_test(wine$pH, wine$quality),
                    simple_cor_test(log10(wine$sulphates), wine$quality),
                    simple_cor_test(wine$alcohol, wine$quality))

names(correlations_2) <- c("fixed.acidity", 
                           "volatile.acidity", 
                           "citric.acid",
                           "log10.residual.sugar",
                           "log10.chlordies", 
                           "free.sulfur.dioxide",
                           "total.sulfur.dioxide", 
                           "density", 
                           "pH",
                           "log10.sulphates", 
                           "alcohol")

correlations_2
```

From the correlation test, it seems that the following variables have a higher correlation to Wine Quality.

1. Alcohol
2. Sulphates(log10)
3. Volatile Acidity
4. Citric Acid

# 7. Summary Analysis of the Bivariate Plots (Summary 2 variables)
### Observations
1. "Fixed Acidity" seems to have almost no effect on quality.
2. "Volatile Acidity" seems to have a negative correlation with the quality.
3. Better wines seem to have higher concentration of "Citric Acid".
4. Better wines seem to have higher "alcohol" percentages. But when I created a linear model around it, I saw from the R squared value that alcohol by itself only contributes like 22 % on the variance of the quality. So there may be some other factors at play here.
5. Even though it’s a weak correlation, but lower percent of "Chloride" seems to produce better quality wines.
6. Better wines seem to have lower "densities". But then again, this may be due to the higher alcohol content in them.
7. Better wines seem to be more "acidic".
8. "Residual sugar" almost has no effect on the wine quality.

# 8. Multivariate Plots
As we saw, that alcohol plays a strong part in the quality of the wine even though it actually contributes only 22 % of the total quality, now we will first make alcohol constant and try to insert a few more variables to see if they contribute to the overall quality in any other way.

### Plot Density, Alcohol and Quality
```{r}
colors <- c("#BCC5CE", "#3E7C59", "#A62B1F", "#F2B138", "#F20587", "#2E038C")
colors

ggplot(data = wine, aes(x = alcohol, y = density, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Density, Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Alcohol") +
  ylab("Density") 
```

### Plot Sulphates, Alcohol and Quality
```{r}
ggplot(data = wine, aes(x = alcohol, y = sulphates, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Sulphates, Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Alcohol") +
  ylab("Sulphates") +
  facet_wrap(~rating)
```

- Wines with higher alcohol content produce better wine if they have higher level of Sulphates.

### Plot Volatile Acidity, Alcohol and Quality
```{r}
ggplot(data = wine, aes(x = alcohol, y = volatile.acidity, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Volatile Acidity, Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Alcohol") +
  ylab("Volatile Acidity") +
  facet_wrap(~rating)
```

- With less concentration of volatile acid and higher concentration of alcohol seems to produce better wines.

### Plot pH, Alcohol and Quality
```{r}
ggplot(data = wine, aes(x = alcohol, y = pH, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("pH, Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Alcohol") +
  ylab("pH") +
  facet_wrap(~rating)
```

- low pH and high Alcohol percentage seems to produce better wines.

### PLot Residual Sugar, Alcohol and Quality
```{r}
ggplot(data = wine, aes(x = alcohol, y = residual.sugar, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Residual Sugar, Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Alcohol") +
  ylab("Residual Sugar") +
  facet_wrap(~rating)
```

- No such correlation between residual sugar and quality.

### PLot Total Sulfur Dioxide, Alcohol and Quality
```{r}
ggplot(data = wine, aes(x = alcohol, y = total.sulfur.dioxide, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Total Sulfur Dioxide, Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Alcohol") +
  ylab("Total Sulfur Dioxide") +
  facet_wrap(~rating)
```

- In general lower Sulphur Dioxide seems to produces better wine even though some high outliers for better wine with high Sulphur Dioxide.

# The effect of Acids on the Quality of Wines.
### Plot Citric Acid, Volatile Acidity and Quality
```{r}
ggplot(data = wine, aes(x = volatile.acidity, y = citric.acid, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Citric Acid, Volatile Acidity and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Volatile Acidity") +
  ylab("Citric Acid") +
  facet_wrap(~rating)
```

- Higher Citric Acid and low Volatile Acid seems to produce better Wines.

### Plot Citric Acid, Fixed Acidity and Quality
```{r}
ggplot(data = wine, aes(x = fixed.acidity, y = citric.acid, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Citric Acid, Fixed Acidity and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Fixed Acidity") +
  ylab("Citric Acid") +
  facet_wrap(~rating)
```

- No much correlation.

### Plot Volatile Acidity, Fixed Acidity and Quality
```{r}
ggplot(data = wine, aes(x = volatile.acidity, y = fixed.acidity, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Volatile Acidity, Fixed Acidity and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Volatile Acidity") +
  ylab("Fixed Acidity") +
  facet_wrap(~rating)
```

- No much correlation.

# 9. Linear Regression Modelling
Now after all these analysis, I am going to take the variables which are most strongly correlated with the quality of the wine and generate a linear model with them.

```{r}
correlation <- cor(wine %>% 
                     select(-X) %>% 
                     select(-rating) %>% 
                     mutate(quality_num = as.numeric(quality)) %>% 
                     select(-quality))

corrplot(corr = correlation,
         number.cex = 0.9, 
         method = "circle", 
         type = "upper",
         tl.cex = 0.8,
         tl.col = "black")
```

### split dataset
```{r}
set.seed(123)
n <- nrow(wine)
sample_data <- sample(1:n, 0.8*n)
train_data <- wine[sample_data, ]
test_data <- wine[-sample_data, ]
```

### Linear Legression between the most correlation variables
```{r}
linear_model_multi <- lm(as.numeric(quality) ~ fixed.acidity + 
                                               volatile.acidity + 
                                               citric.acid +
                                               sulphates + 
                                               alcohol,
                         data = train_data)

summary(linear_model_multi)
```

```{r}
# install.packages("memisc")
library(memisc)

m1 <- lm(as.numeric(quality) ~ alcohol, data = train_data)
m2 <- update(m1, ~ . + sulphates)
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + citric.acid)
m5 <- update(m4, ~ . + fixed.acidity)
m6 <- update(m2, ~ . + pH)
mtable(m1, m2, m3, m4, m5, m6)
```

### Plot Error and Quality
```{r}
wine_predict <- data.frame(test_data$quality, 
                           predict(m5, test_data) - as.numeric(test_data$quality))
names(wine_predict) <- c("quality", "error")

datatable(wine_predict)
```

```{r}
ggplot(data = wine_predict, aes(x = quality, y = error)) + 
  geom_jitter(alpha = 0.4) +
  ggtitle("Error and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Error")
```

# 10. Summary Analysis of the Multivariate Plots (Summary multi variables)
1. High Alcohol and Sulaphate content seems to produce better wines.
2. Citric Acid, even though weakly correlated plays a part in improving the wine quality.

# 11. Final Plots and Summary
1. Alcohol มีความสัมพันธ์เชิงบวกกับ Wine Quality มากที่สุดคิดเป็น 47.62 % รองลงมาเป็น Sulphate(log10) คิดเป็น 30.86 % และ  Citric Acid คิดเป็น 22.64 %
2. Volatile Acidity มีความสัมพันธ์เชิงลบกับ Wine Quality มากที่สุดคิดเป็น 39.06 % 
3. ปริมาณ Alcohol และ Sulphate เพิ่มขึ้น จะทำให้ Wine Quality เพิ่มขึ้น (จาก plot Alchhol, Sulphate and Quality)
4. Volatile Acidity น้อย และ Alcohol มาก จะทำให้ Wine Quality เพิ่มขึ้น
4. จาก plot Linear Regression of Alcohol and Quality ได้ค่า R-squared = 0.2267 หรือคิดเป็น 22.67 % (Alcohol มีความสัมพันธ์กับ Wine Quality 22.67 %) ซึงมีค่าที่ต่ำ เนื่องด้วย Dataset ประกอบด้วย Average Wine Quality เป็นส่วนใหญ่ และมีข้อมูลที่น้อยมากสำหรับ Good and Bad Wine Quality ทำให้คาดเดาข้อมูลทางสถิติยาก และได้ค่าที่ต่ำ

### Plot 1 Alcohol and Quality
```{r}
ggplot(data = wine, aes(x = quality, y = alcohol)) +
  geom_boxplot(color = "blue") +
  geom_jitter(alpha = 0.3) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle("Alcohol and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Alcohol")
```

- Alcohol เป็นส่วนสำคัญต่อ Wine Quality
- Alcohol ที่สูงขึ้น จะทำให้ Wine Quality ดีขึ้น
- ข้อมูลส่วนใหญ่จะเป็น Average Wine Quality 
- จาก Box plot ค่ามัธยฐานที่สูงสุดของ Wine Quality หมายความว่า เกือบทุกจุดมี Alcohol สูง แต่จาก Linear Regression between Alcohol and Quality ค่า R Squared = 22 % (Alocohol มีความสัมพันธ์กับ Quality แค่ 22 %) ดังนั้น Alcohol จึงไม่ใช่ปัจจัยเดียวที่ส่งผลต่อ Wine Quality

### Plot 2 Alcohol and Sulphates over Wine Quality
```{r}
ggplot(data = wine, aes(x = alcohol, y = sulphates, color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_manual(values = colors) +
  ggtitle("Alcohol and Sulphates over Wine Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Sulphate (g/dm3)") +
  xlab("Alcohol") 
```

- Good Wine Quality(7 และ 8) มีปริมาณ Alcohol และ Sulphate ที่สูงร่วมกัน ึแม้เส้นตรงจะมีความลาดเอียงเล็กน้อย อาจเป็นเพราะใน Good Wine Quality มี Alcohol มากกว่า Sulphate เล็กน้อย

### Plot 3 Error and Quality
```{r}
ggplot(data = wine_predict, aes(x = quality, y = error)) + 
  geom_jitter(alpha = 0.4) +
  ggtitle("Error and Quality") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Quality") +
  ylab("Error")
```

- จะเห็นว่าค่า Error ของ Average Wine Quality จะมีความหนาแน่นมากกว่า Good and Bad Wine Quality 
- จะเห็นได้ว่าข้อมูลส่วนใหญ่เป็น Average Wine Quality และไม่มีค่าผิดปกติมากเกินไป (Outliers)
- Linear Regression Model ของ m5 มีค่า R-Squared = 0.335 หรือ 33.5 % หมายความว่า Alcohol, Sulphates, Volatile Acidity, Citric Acid and Fixed Acidity มีความสัมพันธ์กับ Quality 33.5%

# 12. Reference
"https://www.kaggle.com/code/sagarnildass/red-wine-analysis-by-r/report"

"https://www.kaggle.com/code/piyushgoyal443/red-wine-analysis/report#bivariate-plots-section"

"https://r-charts.com/correlation/"

"https://www.interviewbit.com/blog/r-projects/"


