---
title: "Movie Recommendation System"
author: "Chayut"
date: "2022-10-22"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
---

This machine learning project’s main purpose is to create a recommendation engine that can suggest movies to consumers. The goal of this R project is to learn how a recommendation system works. 

# Import Library
```{r}
# install.packages("recommenderlab")
library(recommenderlab)
library(tidyverse)
library(data.table)
library(reshape2)
```

# Import dataset
The MovieLens Dataset was utilized. There are 105339 ratings in the ratings.csv file from which the data was utilized. Ratings were given over 10329 movies in the movies.csv file.

```{r}
Movies_data <- read.csv("movies.csv", stringsAsFactors = FALSE)
Ratings_data <- read.csv("ratings.csv")
```

# Review dataset
```{r}
head(Movies_data)
summary(Movies_data)
```

```{r}
head(Ratings_data)
summary(Ratings_data)
```

# Data pre-processing

# Creating a one-hot encoding to create a matrix that comprises of corresponding genres for each of the films.
```{r}
Movies_genre <- as.data.frame(Movies_data$genres, stringsAsFactors = FALSE)
Movies_genre
```

```{r}
# Split a column using 'tstrsplit'
library(data.table)
Movies_genre_2 <- as.data.frame(tstrsplit(Movies_genre[ , 1], "[|]", 
                                          type.convert = TRUE),
                                stringsAsFactors = FALSE)
Movies_genre_2
```

```{r}
# change name of columns
colnames(Movies_genre_2) <- c(1:10)
Movies_genre_2
```

```{r}
# list genre
list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western") # 18 genre

# create genre matrix
genre_matrix_1 <- matrix(0, nrow = 10330, ncol = 18) # Movies_data 10329 obs.

genre_matrix_1[1, ] <- list_genre

# change name of columns
colnames(genre_matrix_1) <- list_genre

for (index in 1:nrow(Movies_genre_2)) {
  for (col in 1:ncol(Movies_genre_2)) {
    genre_col = which(genre_matrix_1[1, ] == Movies_genre_2[index, col])
    genre_matrix_1[index + 1, genre_col] <- 1
  }
}

# remove first row, which was the genre list
genre_matrix_2 <- as.data.frame(genre_matrix_1[-1, ], stringsAsFactors = FALSE)
genre_matrix_2
```

```{r}
str(genre_matrix_2)
```

```{r}
# convert from characters to integers
for (col in 1:ncol(genre_matrix_2)) {
  genre_matrix_2[ , col] <- as.integer(genre_matrix_2[ , col]) 
}
str(genre_matrix_2)
```

In the next step of Data pre-processing of R project, we will create a ‘search matrix’ that will allow us to perform an easy search of the films by specifying the genre present in our list.

# Creating a ‘search matrix’ - searching films by specifying the genre
```{r}
# combind Movies_data (only colums movieId and title) and genre_matrix_2
Search_Matrix <- cbind(Movies_data[, 1:2], genre_matrix_2[ , ])
head(Search_Matrix)
```

There are movies that have several genres, for example, Toy Story, which is an animated film also falls under the genres of Comedy, Fantasy, and Children. This applies to the majority of the films.

For our movie recommendation system to make sense of our ratings through recommenderlabs, we have to convert our matrix into a sparse matrix one. This new matrix is of the class ‘realRatingMatrix’.

# Convert rating matrix into a recommenderlab sparse matrix
```{r}
names(Ratings_data)
```

```{r}
Ratings_Matrix <- dcast(data = Ratings_data, 
                        formula = userId ~ movieId, # row = userId, col = movieId
                        value.var = "rating", # name of column 
                        na.rm = FALSE)
head(Ratings_Matrix)
```

```{r}
# remove column userID
Ratings_Matrix <- as.matrix(Ratings_Matrix[ , -1])
```

```{r}
str(Ratings_Matrix)
```

```{r}
# Convert rating matrix into a recommenderlab sparse matrix
Ratings_Matrix <- as(object = Ratings_Matrix, Class = "realRatingMatrix")
Ratings_Matrix
```

# Overview some important parameters for building recommendation systems for movies
Let us now overview some of the important parameters that provide us various options for building recommendation systems for movies.

```{r}
# look at registry and a few methods
Recommendation_Model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(Recommendation_Model)
```

```{r}
# 'lapply' returns a List
lapply(Recommendation_Model, "[[", "description") 
```
# Implementing a single model in the R project – Item Based Collaborative Filtering
We will implement a single model in our R project.

'Item Based Collaborative Filtering (IBCF_realRatingMatrix)'.

```{r}
Recommendation_Model$IBCF_realRatingMatrix$parameters
```

# Exploring Similar Data

# Collaborative Filtering involves suggesting movies to the users that are based on collecting preferences from many other users.
# With the help of recommenderlab, we can compute similarities between users
Collaborative Filtering involves suggesting movies to the users that are based on collecting preferences from many other users. For example, if a user A likes to watch action films and so does user B, then the movies that the user B will watch in the future will be recommended to A and vice-versa. Therefore, recommending movies is dependent on creating a relationship of similarity between the two users. With the help of recommenderlab, we can compute similarities using various operators like cosine, pearson as well as jaccard.

```{r}
Similarity_Matrix <- similarity(Ratings_Matrix[1:4, ], # select 4 users
                                method = "cosine",
                                which = "users")
as.matrix(Similarity_Matrix)
```

In the above matrix, each row and column represents a user. We have taken four users and each cell in this matrix represents the similarity that is shared between the two users.

```{r}
image(x = as.matrix(Similarity_Matrix), 
      main = "User's Similarities")
```

In the above matrix, each row and column represents a user. We have taken four users and each cell in this matrix represents the similarity that is shared between the two users.

# Portray the similarity that is shared between the films
Now, we delineate the similarity that is shared between the films.

```{r}
names(Movies_data)
```

```{r}
Movies_Similarity <- similarity(Ratings_Matrix[ , 1:4],
                                method = "cosine",
                                which = "items")
as.matrix(Movies_Similarity)
```

```{r}
image(as.matrix(Movies_Similarity), main = "Movies similarity")
```

Let us now extract the most unique ratings.

```{r}
Ratings_values <- as.vector(Ratings_Matrix@data)
```

```{r}
# extracting unique ratings
unique(Ratings_values)
```

```{r}
# creating a count of movie ratings
Table_of_Ratings <- table(Ratings_values) 
Table_of_Ratings
```

# Most viewed movies visualization
we will explore the most viewed movies in our dataset. We will first count the number of views in a film and then organize them in a table that would group them in descending order.

```{r}
# count views for each movie
Movies_views <- colCounts(Ratings_Matrix)
Movies_views
```

```{r}
# create dataframe of views
table_views <- data.frame(movies = names(Movies_views),
                          views = Movies_views)
head(table_views)
```

```{r}
# sort by number of views
table_views <- table_views[order(table_views$views, decreasing = TRUE), ]
head(table_views)
```

```{r}
# review data
names(table_views)
```

```{r}
summary(table_views)
```

```{r}
# add column title
table_views$title <- NA

# add name title
for (index in 1:10325 ) {
  table_views[index, 3] <- as.character(subset(Movies_data,
                                               Movies_data$movieId == table_views[index, 1])$title)
}

head(table_views)
```

# Most Viewed Movies Visualization
Visualize a bar plot for the total number of views of the top films.

```{r}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat = "identity", fill = "#4169E1") +
  theme_bw() +
  geom_text(aes(label = views), vjust = -0.3, size = 3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Total Views of the Top Films") +
  ylim(c(0,400)) 
```

# Heatmap of Movie Ratings
Visualize a heatmap of the movie ratings

```{r}
image(Ratings_Matrix[1:20, 1:25],
      axes = FALSE,
      main = "Heatmap of the first 25 rows and 25 columns")
```

# Data Preparation
We will conduct data preparation in the following three steps 

1. Selecting useful data.
2. Normalizing data.
3. Binarizing the data.

# 1. Selecting useful data
For finding useful data in our dataset, we have set the threshold for the minimum number of users who have rated a film as 50. This is also same for minimum number of views that are per film. This way, we have filtered a list of watched films from least-watched ones.

```{r}
Ratings_Matrix
```

```{r}
Movies_Ratings <- Ratings_Matrix[rowCounts(Ratings_Matrix) > 50,
                                 colCounts(Ratings_Matrix) > 50]
Movies_Ratings
```

From the above output of ‘movie_ratings’, we observe that there are 420 users and 447 films as opposed to the previous 668 users and 10325 films. We can now delineate our matrix of relevant users as follows.

# describing matrix of relevant users
```{r}
minimum_Movies <- quantile(rowCounts(Movies_Ratings), 0.98)
minimum_Users <- quantile(colCounts(Movies_Ratings), 0.98)

image(Movies_Ratings[rowCounts(Movies_Ratings) > minimum_Movies,
                     colCounts(Movies_Ratings) > minimum_Users],
      main = "Heatmap of the top users and movies")
```

# Visualizing the distribution of the average ratings per user
```{r}
average_Ratings <- rowMeans(Movies_Ratings)
summary(average_Ratings)
```

```{r}
qplot(x = average_Ratings, 
      fill = I("#4169E1"),
      colour = I("black")) +
  ggtitle("Distribution of the average rating per user")
```

# 2. Normalizing data (Data Normalization)
In the case of some users, there can be high ratings or low ratings provided to all of the watched films. This will act as a bias while implementing our model. In order to remove this, we normalize our data. Normalization is a data preparation procedure to standardize the numerical values in a column to a common scale value. This is done in such a way that there is no distortion in the range of values. Normalization transforms the average value of our ratings column to 0. We then plot a heatmap that delineates our normalized ratings.

```{r}
normalized_Ratings <- normalize(Movies_Ratings)
sum(rowMeans(normalized_Ratings) > 0.00001)

image(normalized_Ratings[rowCounts(normalized_Ratings) > minimum_Movies,
                          colCounts(normalized_Ratings) > minimum_Users],
      main = "Normalized Ratings of the Top Users")
```

# 3. Binarizing the data (Data Binarization)
In the final step of our data preparation in this data science project, we will binarize our data. Binarizing the data means that we have two discrete values 1 and 0, which will allow our recommendation systems to work more efficiently. We will define a matrix that will consist of 1 if the rating is above 3 and otherwise it will be 0.

```{r}
binary_minimum_Movies <- quantile(rowCounts(Movies_Ratings), 0.95)
binary_minimum_Users <- quantile(colCounts(Movies_Ratings), 0.95)

good_rated_films <- binarize(Movies_Ratings, minRating = 3)
image(good_rated_films[rowCounts(Movies_Ratings) > binary_minimum_Movies,
                       colCounts(Movies_Ratings) > binary_minimum_Users],
      main = "Heatmap of the top users and movies")
```

# Collaborative Filtering System
We will build this filtering system by splitting the dataset into 80% training set and 20% test set.

# Splitting the dataset into 80% training set and 20% test set
```{r}
# Splitting the dataset into 80% training set and 20% test set
sampled_data <- sample(x = c(TRUE, FALSE),
                       size = nrow(Movies_Ratings),
                       replace = TRUE,
                       prob = c(0.8, 0.2))
training_data <- Movies_Ratings[sampled_data, ]
testing_data <- Movies_Ratings[!sampled_data, ]
```

# Building the Recommendation System
 - k = 30
 
 - use the cosine method 

```{r}
# Building the Recommendation System
recommendation_system <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_system)
```

```{r}
# 'lapply' returns a List
lapply(recommendation_system, "[[", "description")
```

```{r}
recommendation_system$IBCF_realRatingMatrix$parameters
```

# Create a Recommender Model
```{r}
recommen_model <- Recommender(data = training_data,
                              method = "IBCF",
                              parameter = list(k = 30))
recommen_model
class(recommen_model)
```

Using the 'getModel()' function, we will retrieve the recommen_model. We will then find the class and dimensions of our similarity matrix that is contained within model_info. Finally, we will generate a heatmap, that will contain the top 20 items and visualize the similarity shared between them.

```{r}
# Exploring the data science recommendation system model
model_info <- getModel(recommen_model)
class(model_info$sim)
dim(model_info$sim)
top_items <- 20
image(model_info$sim[1:top_items, 
                     1:top_items],
      main = "Heatmap of the first rows and columns")
```

we will carry out the sum of rows and columns with the similarity of the objects above 0. We will visualize the sum of columns through a distribution.

```{r}
# Visualize sum of rows and columns with the similarity of the objects above 0
sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)

sum_cols <- colSums(model_info$sim > 0)

qplot(x = sum_cols, 
      fill = I("#4169E1"), 
      col = I("black")) + 
  ggtitle("Distribution of the column count")
```

# How to build Recommender System on dataset using R?
We will create a top_recommendations variable which will be initialized to 10, specifying the number of films to each user. We will then use the 'predict() function' that will identify similar items and will rank them appropriately. Here, each rating is used as a weight. Each weight is multiplied with related similarities. Finally, everything is added in the end.

```{r}
# the number of items to recommend to each user
top_recommendations <- 10 

predicted_recommendations <- predict(object = recommen_model,
                                     newdata = testing_data,
                                     n = top_recommendations)
predicted_recommendations
```

```{r}
# recommendation for the first user
user1 <- predicted_recommendations@items[[1]] 
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1

for (index in 1:10) {
  movies_user2[index] <- as.character(subset(Movies_data,
                                             Movies_data$movieId == movies_user1[index])$title)
}

movies_user2
```

```{r}
# matrix with the recommendations for each user
recommendation_matrix <- sapply(predicted_recommendations@items,
                                function(x) { as.integer(colnames(Movies_Ratings)[x]) }) 
#dim(recc_matrix)
recommendation_matrix[ , 1:4]
```

```{r}
# Distribution of the Number of Items for IBCF
number_of_items <- factor(table(recommendation_matrix))

qplot(x = number_of_items, 
      fill = I("#4169E1"), 
      col = I("black")) + 
  ggtitle("Distribution of the Number of Items for IBCF")
```

```{r}
number_of_items_sorted <- sort(number_of_items, decreasing = TRUE)
number_of_items_top <- head(number_of_items_sorted, n = 4)
table_top <- data.frame(as.integer(names(number_of_items_top)),
                        number_of_items_top)

for(i in 1:4) {
  table_top[i, 1] <- as.character(subset(Movies_data,
                                        Movies_data$movieId == table_top[i,1])$title)
}

colnames(table_top) <- c("Movie Title", "No. of Items")
head(table_top)
```

# Reference

"https://data-flair.training/blogs/data-science-r-movie-recommendation/"

"https://github.com/Rpita623/Movie-Recommendation-System-using-R_Project"
















