---
title: "Customer Segmentation using K-Means Clustering"
author: "Chayut Phukson"
date: "2022-10-14"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
---

# Import Library
```{r}
library(tidyverse)
library(plotrix) # create pie chart
library(skmeans)
library(gridExtra)
```

# Import Dataset
ref: "https://www.kaggle.com/datasets/vjchoudhary7/customer-segmentation-tutorial-in-python?select=Mall_Customers.csv"
```{r}
Mall_Customers <- read.csv("Mall_Customers.csv", stringsAsFactors = T)
```

```{r}
Mall_Customers
```

# Review Dataset
```{r}
head(Mall_Customers)
```

```{r}
names(Mall_Customers)
```

```{r}
str(Mall_Customers)
```

```{r}
summary(Mall_Customers)
```

```{r}
# check NA
sum(is.na(Mall_Customers))
```

# Customer Gender Visualization
# Using BarPlot to display Gender Comparision
```{r}
Mall_Customers %>% 
  ggplot(aes(Gender)) +
  geom_bar(fill = c("#529cf6", "#ffb855")) +
  ylim(0,120) +
  labs(title = "Gender Comparision") +
  theme_bw() +
  coord_flip() +
  xlab("") 
```

# Pie Chart Depicting Ratio of Female and Male
```{r}
Gender_Table <- table(Mall_Customers$Gender)
Gender_Table
```

```{r}
Percent_Gender <- (table(Mall_Customers$Gender)/sum(table(Mall_Customers))) * 100
Percent_Gender
```

```{r}
Gender_labels <- paste(c("Female","Male"), " ", Percent_Gender, "%", sep = " ")
```

```{r}
# Create pie chart
pie(x = Gender_Table, labels = Gender_labels, col = c("#529cf6", "#ffb855"))
```

# Visualization of Age Distribution
```{r}
table(Mall_Customers$Age)
```

```{r}
hist(Mall_Customers$Age,
    col = "#FFE4B5",
    main = "Histogram to Show Count of Age Class",
    ylim = c(0,40),
    xlab = "Age",
    ylab = "Frequency",
    labels = TRUE)
```

```{r}
Mall_Customers %>% 
  ggplot(aes(x = Age)) +
  geom_boxplot(fill = "#FFE4B5") +
  theme_classic() 
```

# Analysis of the Annual Income of the Customers
# Descriptive Analysis
```{r}
table(Mall_Customers$Annual.Income..k..)
```

```{r}
summary(Mall_Customers$Annual.Income..k..)
```

```{r}
hist(Mall_Customers$Annual.Income..k..,
    col = "#E6E6FA",
    main = "Histogram for Annual Income",
    ylim = c(0, 40),
    xlab = "Annual Income",
    ylab = "Frequency",
    labels = TRUE)
```

```{r}
plot(density(Mall_Customers$Annual.Income..k..),
    main = "Density Plot for Annual Income",
    xlab = "Annual Income",
    ylab = "Density")
polygon(density(Mall_Customers$Annual.Income..k..), col = "#E6E6FA")
```

# Analyzing Spending Score of the Customers
```{r}
table(Mall_Customers$Spending.Score..1.100.)
```

```{r}
summary(Mall_Customers$Spending.Score..1.100.)
```

```{r}
hist(Mall_Customers$Spending.Score..1.100.,
    main = "Histogram for Spending Score",
    ylim = c(0,45),
    xlab = "Spending Score",
    ylab = "Frequency",
    col = "#FFE4E1",
    labels = TRUE)
```

```{r}
Mall_Customers %>% 
  ggplot(aes(x = Spending.Score..1.100.)) +
  geom_boxplot(fill = "#FFE4E1") +
  xlab("Spending Score") +
  theme_classic()
```

# K-means Algorithm
# Elbow Method

Customer Segmentation with Psychographics (social class, lifestyle, personality traits) and Behavioral data (spending and consumption habits,  product/service usage)

# 1. Customer Segmentation with Annual Income and Speding Score

We customer Segmentation with Annual Income of the customee and the Speding Score. 


```{r}
# select variables Annual Income and the Speding Score
k_newdata <- Mall_Customers[ , c(4:5)]
```

```{r}
library(purrr)
set.seed(111)
# function to calculate total within-cluster sum of square (WCSS)
wcss <- function(k) {
  kmeans(k_newdata, 
         k, 
         iter.max = 100, 
         nstart = 100, 
         algorithm = "Lloyd" )$tot.withinss
}
# compute and plot wcss for k = 1 to k = 10
k_values <- 1:10 

wcss_values <- map_dbl(k_values, wcss)

plot(k_values, wcss_values,
     type = "b", 
     pch = 19, 
     frame = FALSE, 
     xlab = "Number of clusters K",
     ylab = "Total within-cluster sum of square")
```

# Compute k-means with k = 5
```{r}
set.seed(111)
customerClusters <- kmeans(k_newdata, 5)
customerClusters
```

# Visualization K-means Cluster of Annual Income and Speding Score
```{r}
set.seed(111)
ggplot(data = k_newdata, aes(x = Annual.Income..k.., y = Spending.Score..1.100.)) + 
  geom_point(stat = "identity", 
             aes(color = as.factor(customerClusters$cluster))) +
  scale_color_discrete(name = " ",
                       breaks = c("1", "2", "3", "4", "5"),
                       labels = c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")) +
  ggtitle("Mall Customer Segmens", subtitle = "Using K-means Clustering") +
  xlab("Annual Income") +
  ylab("Speding Score")
```

Cluster 1. Customers with medium annual income and medium annual spend

Cluster 2. Customers with low annual income and low annual spend 

Cluster 3. Customers with high annual income and high annual spend

Cluster 4. Customers low annual income but high annual spend

Cluster 5. Customers with high annual income and low annual spend


# 2. Customer Segmentation with Age and Spending Score

```{r}
# select variables Age and Spending Score
k_newdata_Age_Spending <- Mall_Customers[ , c(3,5)]
```

```{r}
library(purrr)
set.seed(1)
# function to calculate total within-cluster sum of square (WCSS)
wcss <- function(k) {
  kmeans(k_newdata_Age_Spending, 
         k, 
         iter.max = 100, 
         nstart = 100, 
         algorithm = "Lloyd" )$tot.withinss
}
# compute and plot wcss for k = 1 to k = 10
k_values <- 1:10 

wcss_values <- map_dbl(k_values, wcss)

plot(k_values, wcss_values,
     type = "b", 
     pch = 19, 
     frame = FALSE, 
     xlab = "Number of clusters K",
     ylab = "Total within-cluster sum of square")
```

# Compute k-means with k = 4
```{r}
set.seed(1)
customerClusters_Age_Spending <- kmeans(k_newdata_Age_Spending, 4)
customerClusters_Age_Spending
```

# Visualization K-means Cluster of Age and Speding Score
```{r}
set.seed(1)
ggplot(data = k_newdata_Age_Spending, aes(x = Age, y = Spending.Score..1.100.)) + 
  geom_point(stat = "identity", 
             aes(color = as.factor(customerClusters_Age_Spending$cluster))) +
  scale_color_discrete(name = " ",
                       breaks = c("1", "2", "3", "4"),
                       labels = c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4")) +
  ggtitle("Mall Customer Segmens", subtitle = "Using K-means Clustering") +
  xlab("Age") +
  ylab("Speding Score")
```

Cluster 1. Customers with high age and medium annual spend

Cluster 2. Customers with low age and medium annual spend 

Cluster 3. Customers with low to high of age and low annual spend

Cluster 4. Customers low age but high annual spend


# 3. Customer Segmentation with Age, Anual Income and Spending Score

```{r}
# select variables Age, Anual Income and Spending Score
k_newdata_Age_Income_Spending <- Mall_Customers[ , c(3:5)]
```

```{r}
library(purrr)
set.seed(1)
# function to calculate total within-cluster sum of square (WCSS)
wcss <- function(k) {
  kmeans(k_newdata_Age_Income_Spending, 
         k, 
         iter.max = 100, 
         nstart = 100, 
         algorithm = "Lloyd" )$tot.withinss
}
# compute and plot wcss for k = 1 to k = 10
k_values <- 1:10 

wcss_values <- map_dbl(k_values, wcss)

plot(k_values, wcss_values,
     type = "b", 
     pch = 19, 
     frame = FALSE, 
     xlab = "Number of clusters K",
     ylab = "Total within-cluster sum of square")
```

# Compute k-means with k = 6
```{r}
set.seed(1)
customerClusters_Age_Income_Spending <- kmeans(k_newdata_Age_Income_Spending, 6)
customerClusters_Age_Income_Spending
```


ref: 
"https://data-flair.training/blogs/r-data-science-project-customer-segmentation/"
"https://www.youtube.com/watch?v=Whlis-6vsUY"




# Others fuctions or libraries

```{r}
# select variables that interested
k_newdata2 <- Mall_Customers[ , c(4:5)]
```

```{r}
# K estimation using Elbow approach
set.seed(111)
tot.withinss <- vector("numeric", length = 10)
for (i in 1:10) {
    kDet <- kmeans(k_newdata2, i)
    tot.withinss[i] <- kDet$tot.withinss
}

ggplot(as.data.frame(tot.withinss), aes(x = seq(1,10), y = tot.withinss)) + 
    geom_point(col = "#F8766D") +    
    geom_line(col = "#F8766D") + 
    theme(axis.title.x.bottom = element_blank()) +
    ylab("Within-cluster Sum of Squares") +
    xlab("Number of Clusters") +
    ggtitle("Elbow K Estimation")
```

```{r}
library(factoextra)
fviz_nbclust(k_newdata2, kmeans, method = "wss")
```

```{r}
# Cluster
set.seed(111)
customerClusters2 <- kmeans(k_newdata, 5)
customerClusters2
```

```{r}
# Visualization Cluster
set.seed(111)
ggplot(data = k_newdata2, aes(x = Annual.Income..k.., y = Spending.Score..1.100.)) + 
  geom_point(stat = "identity", 
             aes(color = as.factor(customerClusters2$cluster))) +
  scale_color_discrete(name = " ",
                       breaks = c("1", "2", "3", "4", "5"),
                       labels = c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")) +
  ggtitle("Mall Customer Segmens", subtitle = "Using K-means Clustering") +
  xlab("Annual Income") +
  ylab("Speding Score")
```


Cluster 1. Customers with medium annual income and medium annual spend

Cluster 2. Customers with low annual income and low annual spend 

Cluster 3. Customers with high annual income and high annual spend

Cluster 4. Customers low annual income but high annual spend

Cluster 5. Customers with high annual income and low annual spend

ref: "https://www.kaggle.com/code/ioannismesionis/mall-customer-segmentation-unsupervised-learning/notebook"





























