---
title: "Sentiment Analysis with R"
author: "chayut"
date: "2022-10-10"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
---

- Sentimental analysis is the process of evaluating words to discover sentiments and opinions that may be positive or negative in polarity. 

- With the help of sentiment analysis, you can find out the nature of opinion that is reflected in documents, websites, social media feed, etc. 

- Sentiment Analysis is a type of classification where the data is classified into different classes. 

- This is a type of categorization in which the categories are either binary (optimistic or pessimistic) or multiple (happy, angry, sad, disgusted, etc.).

- The project is written in R, and it makes use of the Janeausten R package’s dataset. 

- To perform an inner join, general-purpose lexicons like AFINN, bing, and Loughran are utilized, and the results are presented using a word cloud.

# Install Package
```{r}
# install.packages("janeaustenr")
```

# Import Library

The dataset that we will use will be provided by the R package ‘janeaustenr’.

```{r}
library(tidyverse)
library(janeaustenr) 
library(DT) 
library(tidytext)
library(stringr)
```

# View Sentiments
```{r}
datatable(sentiments)
```

We will make use of three general purpose lexicons.

1. AFINN

- the AFINN lexicon model scores the words in "a range from -5 to 5".

- The increase in negativity corresponds the negative sentiment whereas an increase in positivity corresponds the positive one.

2. bing

- The bing lexicon model on the other hand, classifies the sentiment into a binary category of "negative or positive".

3. loughran

- The loughran model that performs analysis of "the shareholder’s reports".

# Using of the bing lexicons to extract the sentiments out of our data
```{r}
datatable(get_sentiments("bing"))
```

# R book code [duplicate]

- The "janeaustenr package" will provide us with the textual data in the form of books authored by the novelist Jane Austen. 

- Tidytext will allow us to perform efficient text analysis on our data. 

- We will convert the text of our books into a tidy format using unnest_tokens() function.

# Review The book Sense and Sensibility (austen_books())
```{r}
names(austen_books())
```

```{r}
datatable(unique(austen_books()))
```

# The Jane Austen's Complete Novels
```{r}
# library(janeaustenr)
# library(stringr)
# library(tidytext) 
book_data <- 
  austen_books() %>%
  group_by(book) %>% 
  mutate(line_number = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>% 
  ungroup() %>% 
  unnest_tokens(word, text)

book_data
```

- We have performed the tidy operation on our text such that each row contains a single word.

- We will now make use of the “bing” lexicon to and implement filter() over the words that correspond to joy. 

- We will use "the book Emma" and derive its words to implement out sentiment analysis model.

# Filter positive sentiments of R
```{r}
positive_sentiment <- get_sentiments("bing") %>% 
                        filter(sentiment == "positive")

datatable(positive_sentiment)
```

# Review book data
```{r}
names(book_data)
```

```{r}
summary(book_data)
```

# Count words positive sentiment of the book Emma
```{r}
book_data %>% 
  filter(book == "Emma") %>% # The text of Jane Austen's novel "Emma"
  semi_join(positive_sentiment) %>% 
  count(word, sort = TRUE)
```

In the next step.

- We will use "spread() function" to segregate our data into separate columns of positive and negative sentiments.

- We will then use the "mutate() function" to calculate the total sentiment, that is, the difference between positive and negative sentiment.

```{r}
bing <- get_sentiments("bing") # sentiments of R
Emma_sentiment <- book_data %>% 
                    inner_join(bing) %>% 
                    count(book = "Emma",
                          index = line_number %/% 80,
                          sentiment) %>% # count sentiment
                    spread(key = sentiment, # split between positive and negative
                           value = n, 
                           fill = 0) %>% 
                    mutate(difference_sentiment = positive - negative) # difference between positive and negative 

datatable(Emma_sentiment)
```

# Data Visualization
# visualize Words in the book “Emma” and Difference between Positive and Negative Sentiment
```{r}
# create group column storing either "+" or "-"
Emma_sentiment$group <- cut(Emma_sentiment$difference_sentiment, 
                            c(-Inf, 0, Inf), 
                            labels = c("-", "+"))
```

```{r}
datatable(Emma_sentiment)
```


```{r}
ggplot(data = Emma_sentiment, 
       aes(x = index, 
           y = difference_sentiment, 
           fill = group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("-" = "red", "+" = "limegreen")) +
  ggtitle("Words in the book “Emma” and Difference between Positive and Negative Sentiment") +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  xlab("word in the book Emma") +
  ylim(c(-50,160))
```

# Count the most positive and negative words of the book Emma
```{r}
count_most_word <- book_data %>%
                    inner_join(bing) %>% 
                    count(word, sentiment, sort = TRUE) 

datatable(count_most_word)
```

```{r}
count_most_word <- count_most_word %>% 
                    filter(n > 150) %>% 
                    mutate(n = ifelse(sentiment == "negative", -n, n)) %>% 
                    mutate(word = reorder(word, n))

datatable(count_most_word)
```

# visualize the most positive and negative words of the book Emma
```{r}
ggplot(data = count_most_word, aes(x = word, y = n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  ggtitle("The most positive and negative words of the book Emma") +
  ylab("Sentiment Score") +
  ylim(c(-2000, 2000))
```

# visualization positive and negative words in a single wordcloud
use the "comparision.cloud() function"
```{r}
library(reshape2)
library(wordcloud)

book_data %>% 
  inner_join(bing) %>% 
  count(word, sentiment, sort = TRUE) %>% 
  acast(word ~ sentiment,
        value.var = "n",
        fill = 0) %>% 
  comparison.cloud(colors = c("red", "dark green"),
                   max.words = 100)
```

# Referrences
"https://data-flair.training/blogs/data-science-r-sentiment-analysis-project/"

"https://stackoverflow.com/questions/48731076/regex-divxlc-in-text-analysis-in-r-book-code"

"https://www.interviewbit.com/blog/r-projects/"

"https://www.youtube.com/watch?v=PDObtPGm9X0&list=PLoTScYm9O0GGSiUGzdWbjxIkZqEO-O6qZ&index=71"




