---
title: "Uber Data Analysis Project in R"
author: "chayut"
date: "2022-10-26"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
---

- Companies may use data visualization to better understand complicated datasets and make better decisions.

- We’ll use R packages like ggplot2 to create data analysis in this project.

- We utilize user data to extract insights and provide an accurate prediction of clients who will take Uber trips and rides. 

- The study will look at several criteria such as the number of journeys made in a day, the number of travels made in a month, and so on. 

- As a result of this study, we can determine the average number of passengers that Uber may have in a day, the peak hours when there are more consumers available, the number of trips identified at the highest on which day of the month, and so on.


# Import Library
```{r}
# install.packages("ggthemes")
library(tidyverse)
library(ggthemes) # create extra themes and scales with ggplot2
library(lubridate)
library(DT) # we will be able to interface with the JavaScript Library
library(scales)
```

# Creating vector of colors to be implemented in our plots
```{r}
colors <- c("#CC1011", "#665555", "#05a399", "#cfcaca", "#f5e840", "#0683c9", "#e075b0")
colors
```

# Import Dataset
```{r}
# Read the data for each month separately 
apr_data <- read.csv("uber-raw-data-apr14.csv")
may_data <- read.csv("uber-raw-data-may14.csv")
jun_data <- read.csv("uber-raw-data-jun14.csv")
jul_data <- read.csv("uber-raw-data-jul14.csv")
aug_data <- read.csv("uber-raw-data-aug14.csv")
sep_data <- read.csv("uber-raw-data-sep14.csv")
```

```{r}
# view the data
head(apr_data)
head(may_data)
head(jun_data)
head(jul_data)
head(aug_data)
head(sep_data)
```

```{r}
# Combine the data together 
Uber_data <- rbind(apr_data, may_data, jun_data, jul_data, aug_data, sep_data)
cat("The dimensions of the data are:", dim(Uber_data))
```

```{r}
# view the data
head(Uber_data)
```

we will format the datetime into a more readable format using the Date Time conversion function.

```{r}
Uber_data$Date.Time <- as.POSIXct(Uber_data$Date.Time, format = "%m/%d/%Y %H:%M:%S")
Uber_data$Time <- format(as.POSIXct(Uber_data$Date.Time, 
                                    format = "%m/%d/%Y %H:%M:%S"), 
                         format = "%H:%M:%S")
Uber_data$Date.Time <- ymd_hms(Uber_data$Date.Time)
head(Uber_data)
```

```{r}
Sys.setlocale("LC_TIME", "English")
```

```{r}
# Create individual columns for day, month, year and day of week
Uber_data$day <- factor(day(Uber_data$Date.Time))
Uber_data$month <- factor(month(Uber_data$Date.Time, label = TRUE))
Uber_data$year <- factor(year(Uber_data$Date.Time))
Uber_data$day_of_week <- factor(wday(Uber_data$Date.Time, label = TRUE))
head(Uber_data)
```

```{r}
# Add Time variables as well 
Uber_data$second <- factor(second(hms(Uber_data$Time)))
Uber_data$minute <- factor(minute(hms(Uber_data$Time)))
Uber_data$hour <- factor(hour(hms(Uber_data$Time)))
head(Uber_data)
```

# Data Visualisation
# Plotting the trips by hours in a day
```{r}
hourly_data <- Uber_data %>% 
                group_by(hour) %>% 
                summarise(Total = n())

datatable(hourly_data)
```

```{r}
# Plot the data by hour
ggplot(data = hourly_data, aes(x = hour, y = Total)) +
  geom_bar(stat = "identity",
           fill = "#4169E1",
           color = "black") +
  ggtitle("Trips Every Hour", subtitle = "aggregated today") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma) 

```

# Plotting trips by hour and month
```{r}
# Aggregate the data by month and hour
month_hour_data <- Uber_data %>% 
                    group_by(month, hour) %>% 
                    summarise(Total = n())

datatable(month_hour_data)
```

```{r}
ggplot(data = month_hour_data, aes(x = hour, 
                                   y = Total, 
                                   fill = month)) +
  geom_bar(stat = "identity") +
  ggtitle("Trips by Hour and Month") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
```

# Plotting data by trips during every day of the month
```{r}
# Aggregate data by day of the month 
day_data <- Uber_data %>% 
              group_by(day) %>% 
              summarise(Total = n())

datatable(day_data)
```

```{r}
# Plot the data for the day
ggplot(data = day_data, aes(x = day, y = Total)) +
  geom_bar(stat = "identity",
           fill = "#4169E1",
           color = "black") +
  ggtitle("Trips Every Day") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
```

# Plotting trips by day and month 
```{r}
# Collect data by day and month
day_month_data <- Uber_data %>% 
                    group_by(month, day_of_week) %>% 
                    summarise(Total = n())

datatable(day_month_data)
```

```{r}
ggplot(data = day_month_data, aes(x = month, 
                                  y = Total, 
                                  fill = day_of_week)) +
  geom_bar(stat = "identity", 
           aes(fill = day_of_week), 
           position = "dodge") +
  ggtitle("Trips by Day and Month") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = colors)
```

# Number of Trips place during months in a year
```{r}
month_data <- Uber_data %>% 
                group_by(month) %>% 
                summarise(Total = n())

datatable(month_data)
```

```{r}
ggplot(data = month_data, aes(x = month, 
                              y = Total, 
                              fill = month)) +
  geom_bar(stat = "identity") +
  ggtitle("Trips by month") + 
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma) + 
  scale_fill_manual(values = colors)
```

# Finding out the number of Trips by bases
# Plotting trips by Bases
```{r}
ggplot(data = Uber_data, aes(x = Base)) +
  geom_bar(fill = "#6666CC",
           color = "black") +
  ggtitle("Trips by Bases") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma) +
  ylab("Total")
```

# Plotting trips by Bases and Month
```{r}
ggplot(data = Uber_data, aes(x = Base, fill = month)) + 
  geom_bar(position = "dodge") +
  ggtitle("Trips by Bases and Month") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = colors) +
  ylab("Total")
```

# Plotting trips by Bases and Day of Weak
```{r}
ggplot(data = Uber_data, aes(x = Base, fill = day_of_week)) + 
  geom_bar(position = "dodge") +
  ggtitle("Trips by Bases and Day of Weak") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = colors) +
  ylab("Total")
```

# Heatmap visualization of Day, Hour and Month
# Heatmap by Day and Hour
```{r}
heat_day_hour_data <- Uber_data %>% 
                  group_by(day, hour) %>% 
                  summarise(Total = n())

datatable(heat_day_hour_data)
```

```{r}
ggplot(data = heat_day_hour_data, 
       aes(x = day, 
           y = hour, 
           fill = Total)) +
  geom_tile(color = "white") +
  ggtitle("Heat Map by Day and Hour") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())
```

# Heatmap by Day and Month
```{r}
heat_day_month_data <- Uber_data %>% 
                  group_by(day, month) %>% 
                  summarise(Total = n())

datatable(heat_day_month_data)
```

```{r}
ggplot(data = heat_day_month_data, 
       aes(x = day, 
           y = month, 
           fill = Total)) +
  geom_tile(color = "white") +
  ggtitle("Heat Map by Day and MOnth") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())
```

# Heatmap by Day of Weak and Month 
```{r}
heat_dayofweak_month_data <- Uber_data %>% 
                              group_by(day_of_week, month) %>% 
                              summarise(Total = n())

datatable(heat_dayofweak_month_data)
```

```{r}
ggplot(data = heat_dayofweak_month_data, 
       aes(x = day_of_week,
           y = month, 
           fill = Total)) +
  geom_tile() + # color = ""
  geom_text(aes(x = day_of_week, 
                y = month, 
                label = Total), 
            color = "purple", 
            size = 4) +
  scale_fill_gradient(low = "white", high = "#87CEEB") +
  ggtitle("Heat Map by Day of Weak and Month") +
  xlab("") +
  ylab("") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())
```

# Heatmap by Base and Month
```{r}
heat_base_month_data <- Uber_data %>% 
                    group_by(Base, month) %>% 
                    summarise(Total = n())

datatable(heat_base_month_data)
```

```{r}
ggplot(data = heat_base_month_data, 
       aes(x = Base,
           y = month, 
           fill = Total)) +
  geom_tile() + # color = ""
  geom_text(aes(x = Base, 
                y = month, 
                label = Total), 
            color = "purple", 
            size = 4) +
  scale_fill_gradient(low = "#F5F5DC", high = "#D2B48C") +
  ggtitle("Heat Map by Bases and Month") +
  xlab("") +
  ylab("") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "none")
```

# Heatmap by Base and Day of Weak
```{r}
heat_base_dayofweak_data <- Uber_data %>% 
                              group_by(Base, day_of_week) %>% 
                              summarise(Total = n())
datatable(heat_base_dayofweak_data)
```

```{r}
ggplot(data = heat_base_dayofweak_data, 
       aes(x = Base,
           y = day_of_week, 
           fill = Total)) +
  geom_tile() + # color = ""
  geom_text(aes(x = Base, 
                y = day_of_week, 
                label = Total), 
            color = "purple", 
            size = 4) +
  scale_fill_gradient(low = "white", high = "#f57f17") +
  ggtitle("Heat Map by Bases and Day of Weak") +
  xlab("") +
  ylab("") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "none") +
  xlab("Bases")
```

# Creating a map visualization of rides in New York

```{r}
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004

ggplot(Uber_data, aes(x = Lon, y = Lat)) +
  geom_point(size = 1, color = "blue") +
  scale_x_continuous(limits = c(min_long, max_long)) +
  scale_y_continuous(limits = c(min_lat, max_lat)) +
  theme_map() +
  ggtitle("NYC MAP BASED ON UBER RIDES DURING 2014 (APR-SEP)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(Uber_data, aes(x = Lon, y = Lat, color = Base)) +
  geom_point(size = 1) +
  scale_x_continuous(limits = c(min_long, max_long)) +
  scale_y_continuous(limits = c(min_lat, max_lat)) +
  theme_map() +
  ggtitle("NYC MAP BASED ON UBER RIDES DURING 2014 (APR-SEP) by BASE") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Reference
"https://data-flair.training/blogs/r-data-science-project-uber-data-analysis/"

"https://www.kaggle.com/code/prakharrathi25/uber-data-analysis-in-r/notebook"

"https://www.youtube.com/watch?v=PRdwfszdoo0&list=PLoTScYm9O0GGSiUGzdWbjxIkZqEO-O6qZ&index=77"









